---
title: "ManhattanSalesAnalysis"
author: "Oscar Padilla"
date: "June 6, 2016"
output: 
  html_document:
    keep_md: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache=TRUE)
knitr::opts_chunk$set(echo = TRUE)
```
#MANHATTAN HOUSING DATA ANALYSIS

*CODE Extracted from: O'Neil, Cathy; Schutt, Rachel (2013-10-09). Doing Data Science: Straight Talk from the Frontline*

It requires the packages gdata and plyr

```{r}
require(plyr)
require(gdata)
library(plyr)
library(gdata)
```

Now we will be reading the Excel file downloaded from [The Official Website of the City of New York](http://www1.nyc.gov/home/search/index.page?search-terms=Rolling+sales+update)

```{r}
manhattansales <- read.xls("Data/rollingsales_manhattan.xls", pattern ="BOROUGH")
str(manhattansales)
```

##Data Cleansing

Clean and format the data with regular expressions

```{r}
#convert variable SALE.PRICE from factor into numeric and get rid off of leading digits
manhattansales$SALE.PRICE.N <- as.numeric(gsub("[^[:digit:]]","", manhattansales$SALE.PRICE))
#make all variable names lower case
names(manhattansales) <- tolower(names(manhattansales))
#convert other numeric variables, as well as, date variables
manhattansales $ gross.sqft <- as.numeric(gsub("[^[:digit:]]","", manhattansales $ gross.square.feet))
manhattansales$land.sqft <- as.numeric(gsub("[^[:digit:]]","", manhattansales $ land.square.feet))
manhattansales$sale.date <- as.Date(manhattansales$sale.date)
manhattansales$year.built <- as.numeric(as.character(manhattansales$year.built))
str(manhattansales)
```

##Data Selection

Keep only the actual sales (*manhattansales.sale) and family homes ONLY

```{r}
manhattansales.sale <- manhattansales[manhattansales$sale.price.n!=0,]
manhattansales.homes <- manhattansales.sale[which(grepl("FAMILY", manhattansales.sale$building.class.category)),]
summary(manhattansales.homes)
```

Out of a total of `r nrow(manhattansales)` of the original file, only `r nrow(manhattansales.homes)` are actual sales and family homes

##Outliers Elimination

Remove outliers that seem like they weren't actual sales

```{r}
manhattansales.homes$outliers <- (log10(manhattansales.homes$sale.price.n) <=5) + 0
manhattansales.homes <- manhattansales.homes[which(manhattansales.homes$outliers == 0),]
```

By eliminating log10(sales.price.n) <= 5 we end up with `r nrow(manhattansales.homes)` rows

Finally, let's plot the relation between sales price vs. square footage

```{r}
plot(manhattansales.homes$gross.sqft, manhattansales.homes$sale.price.n)
#Logarithmic Scale
plot(log10(manhattansales.homes$gross.sqft),log10(manhattansales.homes$sale.price.n))
```

#Data Segmentation

Upper East Side ONLY

```{r}
manhattansales.uppereast <- manhattansales.homes[which(grepl("UPPER EAST",manhattansales.homes$neighborhood)),]
plot(manhattansales.homes$gross.sqft, manhattansales.homes$sale.price.n)
plot(log10(manhattansales.homes$gross.sqft), log10(manhattansales.homes$sale.price.n))
```